---
title: "lab-03-eda"
author: "Tyler Gallagher"
format:
   html:
     embed-resources: true
editor: visual
---

## Section 1

```{r}
download.file(
  "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
  destfile = file.path("~", "Downloads", "met_all.gz"),
  method   = "libcurl",
  timeout  = 60
)

met <- data.table::fread(file.path("~", "Downloads", "met_all.gz"))
```

## Section 2

```{r}
dim(met)
head(met)
tail(met)
```

## Section 3

```{r}
str(met)
```

## Section 4

```{r}
table(met$year)
table(met$day)
table(met$hour)
summary(met$temp)
summary(met$elev)
summary(met$wind.sp)
```

```{r}
met[met$elev==9999.0] <- NA
summary(met$elev)
```

```{r}
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)
```

```{R}
summary(met$temp)
```

## Section 5

```{r}
met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)
```

```{r}
summary(met$temp)
```

## Section 6

```{r}
elev <- met[elev==max(elev), ]
summary(elev)
```

```{r}
cor(elev$temp, elev$wind.sp, use="complete")
```

```{r}
cor(elev$temp, elev$hour, use="complete")
```

```{R}
cor(elev$wind.sp, elev$day, use="complete")
```

```{r}
cor(elev$wind.sp, elev$hour, use="complete")
```

```{r}
cor(elev$temp, elev$day, use="complete")
```

## Section 7

```{r}
hist(met$elev, breaks=100)
```

```{r}
hist(met$temp)
```

```{r}
hist(met$wind.sp)
```

```{r}
library(leaflet)
```

```{r}
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>%
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
```

```{r}
library(lubridate)
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)
```

```{r}
elev <- elev[order(date)]
head(elev)
```

```{r}
plot(elev$date, elev$temp, type='l')
```

```{r}
plot(elev$date, elev$wind.sp, type='l')
```

## Section 8

Question: Is there any correlation between wind speed (wind.sp) and elevation (elev)?

```{R}
str(met)

```

```{R}
table(met$wind.sp)
table(met$lat)
```

```{R}
summary(met$wind.sp)
```

```{R}
summary(met$elev)
```

Missing many variables, but okay to proceed. A small amount below 0 in elevation is understandable.

```{R}
hist(met$elev)
hist(met$wind.sp)
```

```{R}
cor(met$elev, met$wind.sp, use="complete")
```

```{R}
mod <- lm(elev ~ wind.sp, data = met)
met[, plot(elev, wind.sp, pch = 19, cex = 0.25, main = "Elevation and Wind Speed", xlab = "Elevation", ylab = "Wind Speed")]
```

There does appear to be some correlation between the variables, though the extent of this and its statistical significance would require more analysis.
