---
title: "lab-05"
author: "Tyler Gallagher"
format:
   html:
     embed-resources: true
editor: visual
---

## Merging the Data

```{r}
library(data.table)
library(dplyr)
library(dtplyr)
```

```{r}
stations <- fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
stations[, USAF := as.integer(USAF)]
# Dealing with NAs and 999999
stations[, USAF   := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY   := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE  := fifelse(STATE == "", NA_character_, STATE)]
# Selecting the three relevant columns, and keeping unique records
stations <- unique(stations[, list(USAF, CTRY, STATE)])
# Dropping NAs
stations <- stations[!is.na(USAF)]
```

```{r}
download.file(
  "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
  destfile = file.path("~", "Downloads", "met_all.gz"),
  method   = "libcurl",
  timeout  = 60
)

met <- data.table::fread(file.path("~", "Downloads", "met_all.gz"))
```

```{r}
head(stations, n = 4)
head(met, n = 4)
nrow(met)
nrow(stations)
```

```{r}
met_stations <- left_join(
  x = met,
  y = stations,
  by = c("USAFID" = "USAF")
)
```

```{r}
nrow(met_stations)
```

```{r}
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]
```

```{r}
met_stations <- left_join(
  x = met,
  y = stations,
  by = c("USAFID" = "USAF")
)
```

```{r}
nrow(met_stations)
```

## Question 1

```{r}
head(met_stations)
```

```{r}
summary(met_stations$temp)
summary(met_stations$wind.sp)
summary(met_stations$atm.press)
```

```{r}
met_stations$temp[met_stations$temp < -16] <- NA
```

```{r}
summary(met_stations$temp)
```

```{r}
met_stations_clean <- na.omit(met_stations)
```

```{r}
USAFID_Median_temp <- met_stations_clean$USAFID[met_stations_clean$temp == 23.5]
print(USAFID_Median_temp)
```

```{r}
median_temp <- 23.5
met_stations$abs_diff <- abs(met_stations$temp - median_temp)
closest_station <- met_stations$USAFID[which.min(met_stations$abs_diff)]
print(closest_station)
```

```{r}
USAFID_Median_wind.sp <- met_stations_clean$USAFID[met_stations_clean$temp == 2.10]
print(USAFID_Median_wind.sp)
```

```{r}
median_wind_sp <- 2.10
met_stations$abs_diff_wind_sp <- abs(met_stations$wind.sp - median_wind_sp)
closest_station_wind_sp <- met_stations$USAFID[which.min(met_stations$abs_diff_wind_sp)]
print(closest_station_wind_sp)
```

```{r}
USAFID_Median_atm.press <- met_stations_clean$USAFID[met_stations_clean$temp == 1014.1]
print(USAFID_Median_atm.press)
```

```{r}
median_atm_press <- 1014.1
met_stations$abs_diff_atm_press <- abs(met_stations$atm.press - median_atm_press)
closest_station_atm_press <- met_stations$USAFID[which.min(met_stations$abs_diff_atm_press)]
print(closest_station_atm_press)
```

```{r}
rows_with_usafid_690150 <- met_stations[met_stations$USAFID == 690150, ]
print(rows_with_usafid_690150$temp)
```

The station with the median temperature is USAFID:720113. The station with both the median wind speed and atmospheric pressure is 690150. I checked to see if the station 690150 has the median temperature as well. It appears that the temperature at this station is more commonly above this median temperature, though it has gotten close multiple times.

## Question 2

```{r}
library(dplyr)
station_distance <- met_stations %>%
  group_by(STATE) %>%
  mutate(distance = sqrt((temp - median(temp))^2 + (wind.sp - median(wind.sp))^2 + (atm.press - median(atm.press))^2)) %>%
  ungroup()
most_representative_stations <- station_distance %>%
  group_by(STATE) %>%
  arrange(distance, lat) %>%
  slice(1) %>%
  ungroup()
print(most_representative_stations)
```

## Question 3

```{r}
library(dplyr)
library(geosphere)
calculate_distance <- function(lat1, lon1, lat2, lon2) {
  if (!is.na(lat1) && !is.na(lon1) && !is.na(lat2) && !is.na(lon2)) {
    dist <- distHaversine(c(lon1, lat1), c(lon2, lat2))
    return(dist)
  } else {
    return(NA)
  }
}
state_midpoints <- met_stations %>%
  group_by(STATE) %>%
  summarize(mid_lat = mean(lat, na.rm = TRUE), mid_lon = mean(lon, na.rm = TRUE))
station_distances <- met_stations %>%
  left_join(state_midpoints, by = "STATE") %>%
  rowwise() %>%
  mutate(distance = calculate_distance(lat, lon, mid_lat, mid_lon))
closest_stations <- station_distances %>%
  group_by(STATE) %>%
  filter(distance == min(distance, na.rm = TRUE)) %>%
  ungroup()
print(closest_stations)
```

```{r}
install.packages("leaflet", repos = "https://cran.cnr.berkeley.edu")
library(leaflet)
map <- leaflet() %>%
    addProviderTiles("OpenStreetMap.Mapnik") %>%
  addCircleMarkers(
    data = closest_stations,
    lat = ~lat,
    lng = ~lon,
    radius = 3, 
    color = "blue",  
    fillColor = "blue", 
    fillOpacity = 0.7,
    popup = ~USAFID,  
    label = ~USAFID  
  ) %>%
  addCircleMarkers(
    data = most_representative_stations,
    lat = ~lat,
    lng = ~lon,
    radius = 6, 
    color = "red",
    fillColor = "red",
    fillOpacity = 0.7,
    popup = ~USAFID,  
    label = ~USAFID  
  )
map
```

## Question 4

```{r}
state_avg_temp <- met_stations %>%
  group_by(STATE) %>%
  summarize(avg_temp = mean(temp, na.rm = TRUE))
state_avg_temp <- state_avg_temp %>%
  mutate(
    classification = case_when(
      avg_temp < 20 ~ "Low",
      avg_temp >= 20 & avg_temp < 25 ~ "Mid",
      avg_temp >= 25 ~ "High",
      TRUE ~ "Unknown"
    )
  )
print(state_avg_temp)
```

```{r}
summary_table <- met_stations %>%
  summarize(
    Number_of_entries = n(),
    Number_of_NA_entries = sum(is.na(temp) | is.na(wind.sp) | is.na(atm.press)),
    Number_of_stations = n_distinct(USAFID),
    Number_of_states_included = n_distinct(STATE),
    Mean_temperature = mean(temp, na.rm = TRUE),
    Mean_wind_speed = mean(wind.sp, na.rm = TRUE),
    Mean_atmospheric_pressure = mean(atm.press, na.rm = TRUE)
  )
print(summary_table)
```
