---
title: "HW_02"
author: "Tyler Gallagher"
format:
   html:
     embed-resources: true
editor: visual
---

## Data Wrangling

```{r}
chs_individual <- read.csv("/Users/TylerGallagher13/Desktop/chs_individual.csv")
chs_regional <- read.csv("/Users/TylerGallagher13/Desktop/chs_regional.csv")
library(dplyr)
```

```{r}
chs <- merge(chs_individual, chs_regional, by = "townname", all.x = TRUE)
```

```{r}
head(chs)
```

```{r}
print(nrow(chs))
print(nrow(chs_individual))
```

There are no duplicates.

```{r}
summary(chs$bmi)
```

```{r}
calculate_mode <- function(x) {
  tbl <- table(x)
  mode_value <- as.numeric(names(tbl[tbl == max(tbl)]))
  if (length(mode_value) == 0) return(NA) else return(mode_value)
}
chs <- chs %>%
  group_by(male, hispanic) %>%
  mutate(bmi = ifelse(is.na(bmi), mean(bmi, na.rm = TRUE), bmi))
chs <- chs %>%
  group_by(male, hispanic) %>%
  mutate(smoke = ifelse(is.na(smoke), calculate_mode(smoke), smoke))
chs <- chs %>%
  group_by(male, hispanic) %>%
  mutate(gasstove = ifelse(is.na(gasstove), calculate_mode(gasstove), gasstove))
chs <- ungroup(chs)
```

```{r}
summary(chs$bmi)
```

```{r}
chs <- chs %>%
  mutate(obesity_level = case_when(
    bmi < 14 ~ "underweight",
    bmi >= 14 & bmi < 22 ~ "normal",
    bmi >= 22 & bmi <= 24 ~ "overweight",
    bmi > 24 ~ "obese",
    TRUE ~ NA_character_
  ))
```

```{r}
head(chs$obesity_level)
```

```{r}
summary_table <- chs %>%
  group_by(obesity_level) %>%
  summarize(
    min_BMI = min(bmi, na.rm = TRUE),
    max_BMI = max(bmi, na.rm = TRUE),
    total_observations = n()
  ) %>%
  ungroup()
print(summary_table)
```

```{r}
chs <- chs %>%
  mutate(smoke_gas_exposure = case_when(
    smoke == 1 & gasstove == 1 ~ "Both Smoke and Gas Stove",
    smoke == 1 & gasstove == 0 ~ "Smoke Only",
    smoke == 0 & gasstove == 1 ~ "Gas Stove Only",
    smoke == 0 & gasstove == 0 ~ "Neither Smoke nor Gas Stove",
    TRUE ~ NA_character_
  ))
```

```{r}
summary_table2 <- chs %>%
  group_by(smoke_gas_exposure) %>%
  summarize(
    count = n()
  ) %>%
  ungroup()
print(summary_table2)
```

```{r}
summary(chs$fev)
```

```{r}
calculate_mode <- function(x) {
  tbl <- table(x)
  mode_value <- as.numeric(names(tbl[tbl == max(tbl)]))
  if (length(mode_value) == 0) return(NA) else return(mode_value)
}
chs <- chs %>%
  group_by(male, hispanic) %>%
  mutate(fev = ifelse(is.na(fev), calculate_mode(fev), fev))
chs <- ungroup(chs)
```

```{r}
summary(chs$fev)
```

## Summary Tables

```{r}
summary_table_townname <- chs %>%
  group_by(townname) %>%
  summarize(
    avg_fev = mean(fev, na.rm = TRUE),
    sd_fev = sd(fev, na.rm = TRUE),
    proportion_asthma = mean(asthma, na.rm = TRUE)
  ) %>%
  ungroup()
print(summary_table_townname)
```

```{r}
summary_table_sex <- chs %>%
  group_by(male) %>%
  summarize(
    avg_fev = mean(fev, na.rm = TRUE),
    sd_fev = sd(fev, na.rm = TRUE),
    proportion_asthma = mean(asthma, na.rm = TRUE)
  ) %>%
  ungroup()
print(summary_table_sex)
```

```{r}
summary_table_obesity_level <- chs %>%
  group_by(obesity_level) %>%
  summarize(
    avg_fev = mean(fev, na.rm = TRUE),
    sd_fev = sd(fev, na.rm = TRUE),
    proportion_asthma = mean(asthma, na.rm = TRUE)
  ) %>%
  ungroup()
print(summary_table_obesity_level)
```

```{r}
summary_table_smoke_gas_exposure <- chs %>%
  group_by(smoke_gas_exposure) %>%
  summarize(
    avg_fev = mean(fev, na.rm = TRUE),
    sd_fev = sd(fev, na.rm = TRUE),
    proportion_asthma = mean(asthma, na.rm = TRUE)
  ) %>%
  ungroup()
print(summary_table_smoke_gas_exposure)
```

## Looking at the Data

1.  What is the association between BMI and FEV (forced expiratory volume)?

```{r}
head(chs$bmi)
head(chs$fev)
tail(chs$bmi)
tail(chs$fev)
```

```{r}
summary(chs$bmi)
summary(chs$fev)
```

```{r}
average_bmi <- mean(chs$bmi, na.rm = TRUE)
print(average_bmi)
average_fev <- mean(chs$fev, na.rm = TRUE)
print(average_fev)
```

```{r}
hist(chs$bmi)
```

```{r}
hist(chs$fev)
```

```{r}
plot(chs$bmi, chs$fev, pch=19, cex=0.5)
```

2.  The plot above appears to show some weak correlation between the variables.

    What is the association between smoke and gas exposure and FEV?

FEV EDA done above.

```{r}
head(chs$smoke_gas_exposure)
tail(chs$smoke_gas_exposure)
```

Distribution of smoke and gas exposure performed above.

```{r}
library(ggplot2)
ggplot(chs, aes(x = smoke_gas_exposure, y = fev)) +
  geom_boxplot() +
  labs(
    title = "Box Plot of FEV by Smoke and Gas Exposure",
    x = "Smoke and Gas Exposure",
    y = "FEV"
  )
```

There is not a very clear association between exposure to smoke and/or gas stove and FEV.

3.  What is the association between PM2.5 exposure and FEV?

FEV EDA done above.

```{r}
head(chs$pm25_mass)
tail(chs$pm25_mass)
```

```{r}
summary(chs$pm25_mass)
```

```{r}
average_pm25_mass <- mean(chs$fev, na.rm = TRUE)
print(average_pm25_mass)
```

```{r}
hist(chs$pm25_mass)
```

```{r}
plot(chs$pm25_mass, chs$fev, pch=19, cex=0.5)
```

There does not appear to be a clear association between PM2.5 and FEV.

## Visualization

1.  Facet plot showing scatterplots with regression lines of BMI vs FEV by \"townname\".

```{r}
ggplot(chs, aes(x = bmi, y = fev)) +
  geom_point() +            
  geom_smooth(method = "lm") +
  labs(
    title = "Scatterplot of BMI vs FEV by Townname",
    x = "BMI",
    y = "FEV"
  ) +
  facet_wrap(~ townname, ncol = 2)
```

2.  Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.

```{r}
library(RColorBrewer)
ggplot(chs, aes(x = fev, fill = obesity_level)) +
  geom_histogram(binwidth = 2.0) +
  labs(
    title = "Stacked Histogram of FEV by Obesity Level",
    x = "FEV",
    y = "Count"
  ) +
  scale_fill_manual(values = c("obese" = "red", "normal" = "blue", "overweight" = "green", "underweight" = "purple"))
```

```{r}
library(RColorBrewer)
ggplot(chs, aes(x = fev, fill = obesity_level)) +
  geom_histogram(binwidth = 2.0, position = "fill") + 
  labs(
    title = "Stacked Histogram of FEV by Obesity Level",
    x = "FEV",
    y = "Count"
  ) +
  scale_fill_manual(values = c("obese" = "red", "normal" = "blue", "overweight" = "green", "underweight" = "purple"))
```

3.  Barchart of BMI by smoke/gas exposure.

```{r}
library(ggplot2)

ggplot(chs, aes(x = smoke_gas_exposure, y = bmi, fill = smoke_gas_exposure)) +
  geom_bar(stat = "identity") +
  stat_summary(
    fun.y = "mean",
    geom = "text",
    aes(label = round(..y.., 2)),
    vjust = -0.5, 
    position = position_dodge(width = 0.9)
  ) +
  labs(
    title = "Bar Chart of BMI by Smoke and Gas Exposure",
    x = "Smoke and Gas Exposure",
    y = "BMI (Average)"
  )

```

Directions unclear; listed average BMI at bottom of each bar.

4.  Statistical summary graphs of FEV by BMI and FEV by smoke/gas exposure category.

```{r}
library(ggplot2)
ggplot(chs, aes(x = bmi, y = fev)) +
  geom_boxplot(fill = "lightblue") +
  labs(
    title = "Statistical Summary of FEV by BMI",
    x = "BMI",
    y = "FEV"
  )
```

```{r}
ggplot(chs, aes(x = smoke_gas_exposure, y = fev, fill = smoke_gas_exposure)) +
  geom_boxplot() +
  labs(
    title = "Statistical Summary of FEV by Smoke and Gas Exposure",
    x = "Smoke and Gas Exposure",
    y = "FEV"
  )
```

5.  A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.

```{r}
library(leaflet)
```

```{r}
map <- leaflet(data = chs) %>%
  addTiles()
map <- map %>% 
  addCircleMarkers(
    lat = ~lat, 
    lng = ~lon, 
    radius = 4, 
    color = "red",
    fillOpacity = 0.7,
    popup = ~paste("Community: ", townname, "<br>PM2.5 Mass: ", pm25_mass, " µg/m³"),
    label = ~townname
  )
map <- map %>% 
  setView(lng = -119.4179, lat = 36.7783, zoom = 6)
map
```

6.  Choose a visualization to examine whether PM2.5 mass is associated with FEV.

```{r}
library(ggplot2)
ggplot(chs, aes(x = pm25_mass, y = fev)) +
  geom_point(color = "blue", alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  labs(
    title = "Scatterplot of PM2.5 Mass vs. FEV with Trendline",
    x = "PM2.5 Mass (µg/m³)",
    y = "FEV"
  )
```

The trendline demonstrates that there may be a very small negative correlation between the two variables.
