---
title: "lab-04"
author: "Tyler Gallagher"
format:
   html:
     embed-resources: true
editor: visual
---

## Section 1

```{R}
library(data.table)
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("met_all.gz")
```

## Section 2

```{r}
library(tidyverse)
```

```{r}
summary(met$temp)
```

```{r}
met <- met[temp>-17]
met2 <- met[order(temp)]
head(met2)
```

```{r}
summary(met$temp)
```

```{r}
met$date = as.Date(paste(met$year, met$month, met$day, sep ="-"))
```

```{r}
data.table::week
```

```{r}
head(met)
table(met$date)
```

```{r}
met.wk1 <- met[day >= 1 & day <= 7]
```

```{r}
head(met.wk1$day)
tail(met.wk1$day)
summary(met.wk1$day)
```

```{r}
met[met$elev==9990.0] <- NA
met.wk1[met.wk1$elev==9999.0] <- NA
summary(met.wk1$temp)
summary(met.wk1$rh)
summary(met.wk1$wind.sp)
summary(met.wk1$vis.dist)
summary(met.wk1$dew.point)
summary(met.wk1$lat)
summary(met.wk1$lon)
summary(met.wk1$elev)
```

```{r}
met.wk1$region <- ifelse(met.wk1$lon < -98.00 & met.wk1$lat >= 39.71, "NW",
                           ifelse(met.wk1$lon < -98.00 & met.wk1$lat <39.71, "SW",
                                  ifelse(met.wk1$lon >= -98.00 & met.wk1$lat >= 39.71, "NE",
                                         ifelse(met.wk1$lon >= -98.00 & met.wk1$lat < 39.71, "SE", NA))))
```

```{r}
table(met.wk1$region)
```

```{r}
met.wk1[, elev_cat := fifelse(elev > 252, "high", "low")]
```

## Section 3

```{r}
help(geom_violin)
```

```{r}
ggplot(met.wk1) + 
  geom_violin(mapping = aes(x = 1, y = temp))
```

I simply plotted out the frequencies of temperature.This is approximately normally distributed temperature.

## Section 4

```{r}
ggplot(data = met.wk1, aes(x = wind.sp, y = dew.point, color = region)) +
  geom_jitter(width = 0.2, height = 0.2, alpha = 0.5) +  # Add jittered points
  geom_smooth(method = "lm", se = FALSE)
```

There is an inconsistent correlation between wind speed and dew point by region. There appears to be a negative association in the southwest but a positive association in all other regions.

## Section 5

```{r}
met.wk1[!is.na(elev_cat)] %>%
  ggplot() + 
  geom_bar(mapping = aes(x = elev_cat, fill = region), position = "dodge")+
  scale_fill_brewer(palette = "Accent") + 
  labs (x = "Elevation Category", y = "n")
```

In relative terms, the Southeast has many more weather stations at low elevation than high. The Northwest, Southwest, and Northeast have more at high elevation category than low elevation category.

## Section 6

```{r}
cran_mirror_url <- "https://cran.cnr.berkeley.edu"
install.packages("package_name", repos = cran_mirror_url)
```

```{r}
install.packages("Hmisc", repos = "https://cran.cnr.berkeley.edu")
library("Hmisc")
l <- met.wk1[!is.na(region)] %>%
  ggplot(mapping = aes (x = region, y = dew.point, fill = region, col = region)) + 
    stat_summary(, fun.data="mean_sdl") + 
    stat_summary (geom = "errorbar")

l
```

Dew point average in each region is seen above.

```{r}
m <- met.wk1[!is.na(region)] %>%
  ggplot(mapping = aes (x = region, y = wind.sp, fill = region, col = region)) + 
    stat_summary(, fun.data="mean_sdl") + 
    stat_summary (geom = "errorbar")

m
```

Wind speed average in each region is seen above.

## Section 7

```{r}
install.packages("leaflet", repos = "https://cran.cnr.berkeley.edu")
```

```{r}
library(leaflet)
```

```{r}
met_avg <- met.wk1[,.(rh = mean(rh,na.rm=TRUE), lat = mean(lat), lon = mean(lon)),  by=c("USAFID")]
met_avg <- met_avg[!is.na(rh)]
```

```{r}
n <- leaflet() %>%
  setView(lng = -95.7129, lat = 37.0902, zoom = 4)
color_pal <- colorNumeric(
  palette = c("blue", "red"),  # Customize the colors here
  domain = met_avg$rh
)
n <- n %>%
  addProviderTiles('OpenStreetMap')
n <- n %>%
  addCircleMarkers(
    data = met_avg,
    lng = ~lon,
    lat = ~lat,
    radius = 5,
    color = ~color_pal(rh)
  )
top_10_humidity <- met_avg %>%
  arrange(desc(rh)) %>%
  head(10)
n <- n %>%
  addMarkers(
    data = top_10_humidity,
    lng = ~lon,
    lat = ~lat,
    label = ~paste("RH:", rh, "%")
  )
legend_values <- seq(min(met_avg$rh), max(met_avg$rh), length.out = 5)
legend_labels <- c("Low", "Moderate", "Moderate-High", "High", "Very High")
n <- n %>%
  addLegend(
    position = "bottomright",
    values = c(10, 30, 50, 70, 90),  
    pal = color_pal,
    title = "Relative Humidity",
    opacity = 1,
    labels = c("Low", "Moderate", "Moderate-High", "High", "Very High")
  )
n
```

The areas of highest relative humidity are mostly concentrated around the coastlines.

## Section 8

```{r}
library(ggplot2)
library(ggpubr)
p <- ggboxplot(data = met.wk1, x = "elev_cat", y = "wind.sp", palette = "jco")
p
```
